---
title: "Salmon-Group20"
author: "Meghan Fletcher, Peyton Moore, Mariano Viz "
date: "2/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(tsibble)

```


```{r}
#INITIAL WRANGLING:

#Read in data and update to lower case:
willamette_fish_passage <- read_csv(here("data", "willamette_fish_passage.csv"))%>% 
  clean_names()

#Mutate date to class "Date" and convert data to time series data frame (tsibble):
willamette_ts <- willamette_fish_passage %>% 
  mutate(date = mdy(date)) %>%
  as_tsibble(key = NULL, index = date)
  


```



## Overview
This “Overview” section (above the tabs) should contain, in whatever order you choose:

● An engaging image (with caption, including photo credit) that is relevant to the dataset

● A brief summary (3 - 4 sentences) of the dataset, and what is included in this “report”

● A map of the fish ladder location (you can make this in R on your own, or include an existing map appropriately licensed, with attribution)

● A professionally formatted data citation

● Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons.



## Fish passage at the Willamette Falls, Willamette River (Oregon) {.tabset}


### Tab 1






### Tab 2






### Tab 3


