---
title: "Salmon-Group20"
author: "Meghan Fletcher, Peyton Moore, Mariano Viz "
date: "2/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(lubridate)
library(tsibble)
library(scales)


```


```{r}
#INITIAL WRANGLING:

#Read in data and update to lower case:
willamette_fish_passage <- read_csv(here("data", "willamette_fish_passage.csv"))%>% 
  clean_names()

#Mutate date to class "Date" and convert data to time series data frame (tsibble):
willamette_ts <- willamette_fish_passage %>% 
  mutate(date = mdy(date)) %>%
  as_tsibble(key = NULL, index = date)
  


```



## Overview
This “Overview” section (above the tabs) should contain, in whatever order you choose:

● An engaging image (with caption, including photo credit) that is relevant to the dataset

● A brief summary (3 - 4 sentences) of the dataset, and what is included in this “report”

● A map of the fish ladder location (you can make this in R on your own, or include an existing map appropriately licensed, with attribution)

● A professionally formatted data citation

● Remember: All code that you used to wrangle the data and prepare the graphs should be available to see if we click on the Code buttons.



## Fish passage at the Willamette Falls, Willamette River (Oregon) {.tabset}


### Tab 1






### Tab 2






### Summary statistics and analysis

```{r}
willamette_fish_annual <- willamette_fish_passage %>%
  mutate(date = mdy(date)) %>%
  mutate(date_year = year(date)) 

coho_annual <- willamette_fish_annual %>%
  count(date_year, wt = coho)
jack_coho_annual <- willamette_fish_annual %>%
  count(date_year, wt = jack_coho)
steelhead_annual <- willamette_fish_annual %>%
  count(date_year, wt = steelhead)

 

ggplot()+ 
  geom_line(data = coho_annual, aes(x = date_year, y = n, color = "Coho"), size = 0.7) +
  geom_line(data = jack_coho_annual, aes(x = date_year, y = n, color = "Jack Coho"), size = 0.7) +
  geom_line(data = steelhead_annual, aes(x = date_year, y = n, color = "Steelhead"), size = 0.7) +
  scale_x_continuous(breaks=c(2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010), limits = c(2001,2010)) + 
  scale_color_manual(values = c( "Coho" = "red3", "Jack Coho" = "orange1", "Steelhead" = "palegreen4"))+
  labs(color = "Species")+
  scale_y_continuous(labels = scales::label_number_si())+
  labs(x = "Year",
       y = "Counts",
       title = "Annual counts by Species, Willamette Falls (2001-2010)") +
  theme_minimal()



```

**Figure 1.** Annual counts by species at the Willamette Falls fish ladder, Willamette
River, Oregon (2001-2010). Red, orange, and green lines indicate counts for the species coho, jack coho, and steelhead, respectively. Data: [Columbia River DART](http://www.cbr.washington.edu/dart/query/adult_graph_text) (Data Access in Real Time).




**Summary**

Annual counts of coho, jack coho, and steelhead species on the Willamette Falls fish ladder (2001-2010) show the following major trends:  

● An overall low number of the jack coho species for the entire time period 2001-2010 

● A marked increase in the counts of individuals of the coho species for the years 2009-2010

● A high number of individuals of the steelhead species for the period 2001-2004 followed by a relative decrease in the years 2005-2009 and a recent increase in the year 2010





